---
title: "Calculating conversion tables"
author: "Cornelius Erfort"
date: "5/17/2021"
output: 
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, tidy.opts=list(width.cutoff = 80), tidy = T, python.reticulate = F)
```


# Setting up

This script requires the shape files and election results data which are not included on GitHub. All files except the shapefiles until 2009 are available for download online. See the below for more details.

## Loading packages

```{r packages, message=FALSE, warning=FALSE, results='hide'}

packages <- c(
  "stringr", "readxl", "dplyr", "tidyr", "plyr", "httr", "rvest", "urltools", "XML", "gnumeric", "english", "gsubfn", "zoo", "haven", "xlsx", "lubridate", "ggplot2", "maptools", "sp", "rgdal", "tools", "readr", "spatstat", "rgeos", "corrgram")

lapply(packages[!(packages %in% rownames(installed.packages()))], install.packages)

invisible(lapply(packages, require, character.only = T))

```

# 2017

## Load 2017 county-level geodata

In a first step, the shapefiles and election results for 2017 are loaded. The years until 1953 will be converted to 2017 counties.

```{r 2017-geo}
# Read shapefile
counties_17 <- readOGR(dsn = "data/county shapefiles/2013+2017/Kreisgrenzen_2017_mit_Einwohnerzahl-shp", layer = "Kreisgrenzen_2017_mit_Einwohnerzahl", stringsAsFactors = FALSE) %>% spTransform(CRS("+proj=longlat +datum=WGS84")) 

# Remove East Germany (including Berlin)
counties_17 <- counties_17[counties_17$AGS < 11000, ] 

# Plot map
plot(counties_17)

# Add rownumber
counties_17@data$rownumber <- row(counties_17@data)[, 1]
```

## Load 2017 voting results

```{r 2017-election}
# Read election results
results_17 <- read.csv2("data/Municipality-level results/btw17_wbz/btw17_wbz_zweitstimmen.csv", skip = 4, fileEncoding = "latin1") %>% 
  filter(Land < 11) # Remove East Germany

# Generate a county key
results_17$AGS <- str_c(str_pad(results_17$Land, 2, pad = "0"), results_17$Regierungsbezirk, str_pad(results_17$Kreis, 2, pad = "0"))

# Only keep relevant vars
results_17 <- select(results_17, c(Land, Gültige, Wahlberechtigte..A., CDU, CSU, SPD, DIE.LINKE, GRÜNE, FDP, AfD, AGS)) 

# Aggregate by county
results_17 <- aggregate(. ~ Land + AGS, data = results_17, FUN = sum)

# Combine CDU.CSU (and remove individual columns)
results_17$CDU.CSU <- results_17$CDU + results_17$CSU
results_17 <- select(results_17, -c(CDU, CSU)) %>% dplyr::rename(state = Land, valid_votes = Gültige, reg_voters = Wahlberechtigte..A.)

# Add year suffix to varnames
names(results_17)[-c(1:2)] <- str_c(names(results_17)[-c(1:2)], ".17")

# Make vote shares (party votes divided by valid votes)
results_17[, names(results_17) %>% str_subset("(CDU)|(CSU)|(SPD)|(FDP)|(LINKE)|(GRÜNE)|(AfD)")] <- 
  apply(results_17[, names(results_17) %>%
                     str_subset("(CDU)|(CSU)|(SPD)|(FDP)|(LINKE)|(GRÜNE)|(AfD)")], MARGIN = 2, FUN = function (x) x / results_17[, names(results_17) %>% str_subset("valid_votes")])

```

# Calculate intersections with 2017

## County level: 1953-1976

### Geodata: 1953-1976
```{r 53-76-geo}
county_years <- c(1953, 1957, 1961, 1965, 1969, 1972, 1976)

for(year in county_years) {
  print(year)
  
  # Get 2-digit year
  (shyear <- substr(year, 3, 4) %>% as.numeric()) 
  
  # Load county-level geodata for that year
  counties_ <- readOGR(dsn = "data/county shapefiles/1949-1989/Federal Republic of Germany and German Democratic Republic 1949-1989", layer = str_c("Germany_", year,"_v.1.0"), stringsAsFactors = FALSE, verbose = F) %>% spTransform(CRS("+proj=longlat +datum=WGS84"))
  
  # Correct duplicate counties (usually city and surroundings, e.g. Augsburg and Augsburg (Stadt))
  for (county in counties_@data$GEN[duplicated(counties_@data$GEN)]) {
    # cat(county, ", ")
    counties_@data$GEN[counties_@data$AREA == min(counties_@data$AREA[counties_@data$GEN == county]) & counties_@data$GEN == county] <- counties_@data$GEN[counties_@data$AREA == min(counties_@data$AREA[counties_@data$GEN == county]) & counties_@data$GEN == county] %>% str_c(., ", Stadt")
  }
  
  # Remove East Germany (including Berlin)
  counties_ <- counties_[counties_$KREIS_KENN < 11001000, ] 
  rownames(counties_@data) <- NULL
  
  # Remove Saarland for 1953 (not part of Germany at the time)
  if(year == 1953) counties_ <- counties_[(counties_@data$KREIS_ID %>% substr(1,2)) != "10", ]
  
  # Add rownumber
  counties_@data$rownumber <- row(counties_@data)[, 1]
  
  # Find intersections between year and 2017
  intersections <- gIntersects(counties_17,counties_, byid = TRUE) # Matrix that indicates intersections -->  gOverlaps gives the same results.
  dim(intersections) # Dimensions (height x width)
  
  # Calculate intersection shares between year and 2017
  intersection_shares <- matrix(nrow = nrow(intersections), ncol = ncol(intersections)) # Create matrix for intersection shares (how much of polygon A lies in polygon B)
  # Go through all intersections and calculate areas
  if(!file.exists(str_c("conversion tables/intersection_shares_", shyear,".RData"))) { # Load if file already exists
    for (new_shape in 1:ncol(intersections)) { # Go through all rows
      # cat("o") # Report progess
      for(old_shape in which(intersections[, new_shape])) { # Go through all intersecting columns
        # cat(".") # Report progess
        
        intersection <- gIntersection(counties_17[new_shape, ],  gBuffer(counties_[old_shape, ], byid=TRUE, width=0), checkValidity = T)
        
        # Area of intersection between "old" and 2017 counties divided by area of "old" counties
        intersection_shares[old_shape, new_shape] <- gArea(gIntersection(counties_17[new_shape, ], gBuffer(counties_[old_shape, ], byid=TRUE, width=0), checkValidity = T)) / gArea(counties_[old_shape, ]) 
      }
    }
    
    # Save intersection_shares
    save(intersection_shares, file = str_c("conversion tables/intersection_shares_", shyear,".RData"))
    
  } else load(str_c("conversion tables/intersection_shares_", shyear,".RData"))
  
  # Create object for geodata for that year
  assign(str_c("counties_", shyear), counties_)
  cat("\n\n")
}

```

Correct data with hand coded lists

```{r 53-76-geo-correct, include=FALSE}
shapenames <- read.xlsx("county names/53-76shapenames.xlsx", sheetIndex = 1) %>%
  filter(!is.na(year)) %>%
  suppressWarnings()
for (i in 1:nrow(shapenames)) {
  print(shapenames[i, ])
  counties_ <- get(str_c("counties_", substr(shapenames$year[i], 3, 4) %>% as.numeric())) # Load county SPDF
  counties_@data$GEN[counties_@data$GEN == shapenames$GEN1[i]] <- shapenames$GEN2[i] # Make change
  assign(str_c("counties_", substr(shapenames$year[i], 3, 4) %>% as.numeric()), counties_) # Write county SPDF
}
```

### Election data: 1953-1976

```{r 53-76-election}

vote <- data.frame()
counties_GEN <- data.frame()

for(year in county_years) {
  print(year)
  
   # 2-digit year
  (shyear <- substr(year, 3, 4) %>% as.numeric())
  
  # Load voting results for that year
  vote <- read.csv2(str_c("data/County-level results/btw", year,"kreis.csv"), skip = ifelse(shyear %in% c(57, 61, 65), 5, 4), fileEncoding = ifelse(shyear == 76, "latin1", "UTF-8")) %>% filter(!is.na(Lfd..Nr.)) %>%
    select(c(Kreisfreie.Stadt.bzw..Stadtkreis..Landkreis.bzw..Kreis, Gültige, Wahlberechtigte, CDU, SPD, F.D.P., CSU)) %>% dplyr::rename(FDP = F.D.P.) %>% mutate(year = year) %>% rbind.fill(vote) # Select and rename vars
  
  counties_GEN <- rbind.fill(counties_GEN, data.frame(GEN = get(str_c("counties_", shyear))@data$GEN, year = year))
  
}


```

Correct data with hand coded lists because some names do not match the geodata.

```{r 53-76-election-correct}

# Correct data
vote$Kreisfreie.Stadt.bzw..Stadtkreis..Landkreis.bzw..Kreis[vote$Kreisfreie.Stadt.bzw..Stadtkreis..Landkreis.bzw..Kreis == "Friedberg" & vote$CSU > 0] <- "Friedberg, Stadt" # There are two counties with this name (one is a city county in Bavaria)

vote$Kreisfreie.Stadt.bzw..Stadtkreis..Landkreis.bzw..Kreis[vote$Kreisfreie.Stadt.bzw..Stadtkreis..Landkreis.bzw..Kreis == "Konstanz, Stadt"] <- "Konstanz" # Konstanz is only one county in the geodata
vote$Kreisfreie.Stadt.bzw..Stadtkreis..Landkreis.bzw..Kreis[vote$Kreisfreie.Stadt.bzw..Stadtkreis..Landkreis.bzw..Kreis == "Lüdenscheid, Stadt"] <- "Lüdenscheid" # Lüdenscheid is only one county in the geodata
vote <- aggregate(. ~ Kreisfreie.Stadt.bzw..Stadtkreis..Landkreis.bzw..Kreis + year, data = vote, FUN = sum)

vote$Kreisfreie.Stadt.bzw..Stadtkreis..Landkreis.bzw..Kreis[str_detect(vote$Kreisfreie.Stadt.bzw..Stadtkreis..Landkreis.bzw..Kreis, "(Neuss)|(Neuß)]")] <- "Neuss"

votenames <- read.xlsx("county names/53-76votenames.xlsx", sheetIndex = 1)
for(i in 1:nrow(votenames))  vote$Kreisfreie.Stadt.bzw..Stadtkreis..Landkreis.bzw..Kreis[vote$Kreisfreie.Stadt.bzw..Stadtkreis..Landkreis.bzw..Krei == votenames$name1[i]] <- votenames$name2[i]

# Combine CDU.CSU
vote <- mutate(vote, CDU.CSU = CDU + CSU) %>% select(-c(CDU, CSU)) %>% dplyr::rename(valid_votes = Gültige, reg_voters = Wahlberechtigte)

# Export for manual matching
table(vote$year)
table(counties_GEN$year)
vote <- vote[!(vote$Kreisfreie.Stadt.bzw..Stadtkreis..Landkreis.bzw..Kreis == "Donaueschingen" & vote$year == 1972), ]

for(year in county_years) {
  print(year)
  
  (shyear <- substr(year, 3, 4) %>% as.numeric()) # 2-digit year
  code_GEN <- counties_GEN$GEN[counties_GEN$year == year] %>% unique %>% sort 
  code_VOTE <- vote$Kreisfreie.Stadt.bzw..Stadtkreis..Landkreis.bzw..Kreis[vote$year == year] %>% unique %>% sort
  
  data.frame(GEN = code_GEN, VOTE = code_VOTE) %>% mutate(check = !(GEN == VOTE)) %>% write.xlsx(str_c("county names/", shyear, "_kreis2shapefile.xlsx"))
  
}

for(year in county_years) {
  print(year)
  
  (shyear <- substr(year, 3, 4) %>% as.numeric()) # 2-digit year
  code_GEN <- counties_GEN$GEN[counties_GEN$year == year] %>% unique %>% sort 
  code_VOTE <- vote$Kreisfreie.Stadt.bzw..Stadtkreis..Landkreis.bzw..Kreis[vote$year == year] %>% unique %>% sort
  
  data.frame(GEN = code_GEN, VOTE = code_VOTE) %>% mutate(check = !(GEN == VOTE)) %>% write.xlsx(str_c("county names/", shyear, "_kreis2shapefile.xlsx"))
  
}

```
### Project the 1953-1976 election results onto 2017

```{r 53-76-project}
# Start all results dataframe with 2017 results
results <- results_17

# Go through years
for(year in county_years) {
  print(year)

  (shyear <- substr(year, 3, 4) %>% as.numeric()) # 2-digit year
  
  # Load geodata and vote results for that year
  counties_ <- get(str_c("counties_", shyear))
  vote_ <- vote[vote$year == year, ]
  
  # Add geodata county names to results
  vote_ <- merge(vote_, read.xlsx(str_c("county names/", shyear, "_kreis2shapefile_.xlsx"), sheetIndex = 1, colIndex = 2:3), by.x = "Kreisfreie.Stadt.bzw..Stadtkreis..Landkreis.bzw..Kreis", by.y = "VOTE", all.x = T)
  
  # Add results to geodata
  counties_@data <- merge(counties_@data, vote_, by = "GEN", all.x = T)
  
  # Bring in original order
  counties_ <- counties_[order(counties_@data$rownumber), ]
  
  # Load intersection shares
  load(str_c("conversion tables/intersection_shares_", shyear,".RData"))
  
  # Construct hypothetical results by 2017 counties
  t_results <- data.frame(AGS = counties_17@data$AGS,
                             valid_votes = colSums(counties_@data$valid_votes * intersection_shares, na.rm = T),
                             reg_voters = colSums(counties_@data$reg_voters * intersection_shares, na.rm = T),
                             CDU.CSU = colSums(counties_@data$CDU.CSU * intersection_shares, na.rm = T),
                             SPD = colSums(counties_@data$SPD * intersection_shares, na.rm = T),
                             FDP = colSums(counties_@data$FDP * intersection_shares, na.rm = T))
  
  # Write year in varnames
  names(t_results)[-1] <- str_c(names(t_results)[-1], ".", shyear)
  
  # Make vote shares
  t_results[, names(t_results) %>% str_subset("(CDU)|(CSU)|(SPD)|(FDP)")] <- apply(t_results[, names(t_results) %>% str_subset("(CDU)|(CSU)|(SPD)|(FDP)")], MARGIN = 2, FUN = function (x) x / t_results[, names(t_results) %>% str_subset("valid_votes")])

  # Add this year to all results dataframe  
  results <- merge(results, t_results, by = "AGS", all = T)

}
  
# Set 1953 results for Saarland to NA
results[results$state == "10", names(results) %>% str_subset("\\.53")] <- NA
```

## Municipality level: 1980-2013
```{r 80-13-geo}
munip_years <- c(1980, 1983, 1987, 1990, 1994, 1998, 2002, 2005, 2009, 2013)


```

# Sources

## Shapefiles

**1953-2009**
MPIDR [Max Planck Institute for Demographic Research] and CGG [Chair for Geodesy and Geoinformatics, University of Rostock] 2011: MPIDR Population History GIS Collection (partly based on Bundesamt für Kartographie und Geodäsie 2011) – Rostock.

Bundesamt für Kartographie und Geodäsie 2011: VG 2500 Verwaltungsgebiete 1:2 500 000. Stand 01.01.2009 – Frankfurt am Main.

**2013**
GeoBasis-DE / Bundesamt für Kartographie und Geodäsie (BKG) 2013: Verwaltungsgebiete 1 : 250 000 VG250.
https://hub.arcgis.com/datasets/esri-de-content::kreisgrenzen-2013?geometry=-20.088%2C46.270%2C40.996%2C55.886

**2017**
GeoBasis-DE / Bundesamt für Kartographie und Geodäsie (BKG) 2017: Verwaltungsgebiete 1 : 250 000 VG250.
https://hub.arcgis.com/datasets/b2e6d8854d9744ca88144d30bef06a76_1?geometry=-20.088%2C46.270%2C40.996%2C55.886

**2020**
GeoBasis-DE / Bundesamt für Kartographie und Geodäsie (BKG) 2020: Verwaltungsgebiete 1 : 250 000 VG250.
License: dl-de/by-2-0 (www.govdata.de/dl-de/by-2-0)
https://gdz.bkg.bund.de/index.php/default/digitale-geodaten/verwaltungsgebiete/verwaltungsgebiete-1-250-000-ebenen-stand-01-01-vg250-ebenen-01-01.html

## Election results
### County level

**1953**
Der Bundeswahlleiter 1953: Bundestagswahl 1953, Ergebnisse nach kreisfreien Städten und Landkreisen.
https://www.bundeswahlleiter.de/dam/jcr/9d55fc40-22a5-43f2-9b50-0bd3ff3780fd/btw1953kreis.csv

**1957**
Der Bundeswahlleiter 1957: Bundestagswahl 1957, Ergebnisse nach kreisfreien Städten und Landkreisen.
https://www.bundeswahlleiter.de/dam/jcr/4c84aa1b-4b75-43c5-bfc7-da240208451b/btw1957kreis.csv

**1961**
Der Bundeswahlleiter 1961: Bundestagswahl 1961, Ergebnisse nach kreisfreien Städten und Landkreisen.
https://www.bundeswahlleiter.de/dam/jcr/d3927d97-7c88-4d9a-a08d-97f68a8ce9af/btw1961kreis.csv

**1965**
Der Bundeswahlleiter 1965: Bundestagswahl 1965, Ergebnisse nach kreisfreien Städten und Landkreisen.
https://www.bundeswahlleiter.de/dam/jcr/6b2137a1-01e1-481e-a221-e4de19c588bd/btw1965kreis.csv

**1969**
Der Bundeswahlleiter 1969: Bundestagswahl 1969, Ergebnisse nach kreisfreien Städten und Landkreisen.
https://www.bundeswahlleiter.de/dam/jcr/c5819a55-5a24-4f09-9845-d8f40da89208/btw1969kreis.csv

**1972**
Der Bundeswahlleiter 1972: Bundestagswahl 1972, Ergebnisse nach kreisfreien Städten und Landkreisen.
https://www.bundeswahlleiter.de/dam/jcr/3a7bbc36-c3ff-4a07-aa7c-aec1f3aa05d6/btw1972kreis.csv

**1976**
Der Bundeswahlleiter 1976: Bundestagswahl 1976, Ergebnisse nach kreisfreien Städten und Landkreisen.
https://www.bundeswahlleiter.de/dam/jcr/91e7563c-5149-42c6-8c48-e88390962cdc/btw1976kreis.csv

### Municipality level

**1980**
Der Bundeswahlleiter 1980: Bundestagswahl 1980, Ergebnisse nach Gemeinden
https://www.bundeswahlleiter.de/dam/jcr/bd5c49db-8d83-463b-bab4-630b418b99cf/btw80_wbz.zip

**1983**
Der Bundeswahlleiter 1983: Bundestagswahl 1983, Ergebnisse nach Wahlbezirken und Gemeinden.
https://www.bundeswahlleiter.de/dam/jcr/8b384f87-bad0-4b99-a7a9-88af7f5f01d3/btw83_wbz.zip

**1987**
Der Bundeswahlleiter 1987: Bundestagswahl 1987, Ergebnisse nach Wahlbezirken.
https://www.bundeswahlleiter.de/dam/jcr/7590fb53-3965-438b-a339-d63878bf0e64/btw87_wbz.zip

**1990**
Der Bundeswahlleiter 1990: Bundestagswahl 1990, Ergebnisse nach Wahlbezirken und Gemeinden.
https://www.bundeswahlleiter.de/dam/jcr/f57480e9-f383-4782-9488-fae3d206a84c/btw90_wbz.zip

**1994**
Der Bundeswahlleiter 1994: Bundestagswahl 1994, Ergebnisse nach Wahlbezirken und Gemeinden.
https://www.bundeswahlleiter.de/dam/jcr/538a1f46-4407-49b4-995b-02bd9e3a8399/btw94_wbz.zip

**1998**
Der Bundeswahlleiter 1998: Bundestagswahl 1998, Ergebnisse nach Wahlbezirken und Gemeinden.
https://www.bundeswahlleiter.de/dam/jcr/e2063bc8-4827-4826-a2db-a6cae5d9cf15/btw98_wbz.zip

**2002**
Der Bundeswahlleiter 2002: Bundestagswahl 2002, Ergebnisse nach Wahlbezirken.
https://www.bundeswahlleiter.de/dam/jcr/16a4f1b0-fbd6-4c79-b5a6-b2792ba6df72/btw02_wbz.zip

**2005**
Der Bundeswahlleiter 2005: Bundestagswahl 2005, Ergebnisse nach Wahlbezirken.
https://www.bundeswahlleiter.de/dam/jcr/3a8aa53f-53a1-4ad9-938f-0aa52c18863d/btw05_wbz.zip

**2009**
Der Bundeswahlleiter 2009: Bundestagswahl 2009, Ergebnisse nach Wahlbezirken.
https://www.bundeswahlleiter.de/dam/jcr/159344fc-d466-4d34-b432-25d7891ab09f/btw09_wbz.zip

**2013**
Der Bundeswahlleiter 2013: Bundestagswahl 2013, Ergebnisse nach Wahlbezirken.
https://www.bundeswahlleiter.de/dam/jcr/0ad35576-0c4b-4fa5-85f5-284618b8fa25/btw13_wbz.zip

**2017**
Der Bundeswahlleiter 2017: Bundestagswahl 2017, Ergebnisse nach Wahlbezirken.
https://www.bundeswahlleiter.de/dam/jcr/a2eef6bd-0225-447c-9943-7af0f46c94d1/btw17_wbz.zip

## For details on redistricting see:

**before 1990**
https://de.wikipedia.org/wiki/Kreisreformen_in_der_Bundesrepublik_Deutschland_bis_1990

**after 1990** 
https://de.wikipedia.org/wiki/Kreisreformen_in_Deutschland_nach_1990
